\section{bigdata university}
MA4125 : Data Analysis Lecture 4a
  
    640K ought to be enough for anybody. (Attributed to Bill Gates, 1981)
 
Handouts (please retain):
 
1) The Data Deluge - The Economist Special report. 27th February 2010. 
 
2) "For Today's Graduate, Just One Word: Statistics"
    
    by Steve Lohr, New York Times, August 5th 2009. 
 
3) 10 Hottest Careers in America.
 
    by Lynn O'Shaughnessy CBS Moneywatch.
 
This article is based on UC San Diego Extension's special report "Hot Careers for College graduates 2010". The first three careers are of particular relevance, as data analysis is a significant part of each.

Class
In class discussion of following topics (further to handouts).
 
1) Role of data analysis in the contemporary economy.
 
 
 
2) Evolution of the role of the Data Analyst and Statisician.
 
 
 
3) The Internet as a data collection mechanism.
 
 
 
4) Acceleration of data creation a.k.a "The Data Deluge".
 
 
 
5) The Relationship between Data and Knowledge.
 
 
 
6) Analytics and business intelligence.
 
 
 
7) "Big Data". (What is "Big Data"? What are the 6 principles of "Big Data"?.)
 
 
 
8) The importance of metadata (i.e. information about information.)
\section{MariaDB}

\section{Education at 10gen}
education.10gen.com
This course will go over basic installation, JSON, schema design, querying, insertion of data, indexing and working with language drivers. We will also cover working in sharded and replicated environments. In the course, you will build a blogging platform, backed by MongoDB. Our code examples will be given in Python. A brief Python introduction is included in the course. 


Several worldwide reports have highlighted a shortage of talent in this area. 
The McKinsey Global Institute (2011) has reported that there will be a shortage 
of talent necessary for organisations to take advantage of business intelligence 
and data analytics. MGI predicts that “by 2018, the United States 
alone could face a shortage of 140,000 to 190,000 people with deep 
analytical skills as well as 1.5 million managers and analysts with
 the know-how to use the analysis of big data to make effective decisions”. 
A 2011 Bloomberg Businessweek Research Services survey of global businesses 
(published in 2012) shows that 97% of companies with revenues of more than 
$100 million are using some form of business analytics. 
43\% of these companies reported plans to increase analytical talent. Locally some estimates indicate a skills shortage of over 3,000 graduates with analytics skills in Ireland during the same period. This programme has been developed in consultation with large organisations in IT, finance and consulting in order to suit their business needs and thus securing future employment for graduates. 
%-----------------------------%
\section{Mortar Data}

Mortar Data is Hadoop for developers, plain and simple. The company has offered its cloud service — which replaces MapReduce with a combination Pig and Python — for almost a year. In November, it released the open source Mortar framework in order to build a community around sharing datasets and making it easier to write Hadoop pipelines. Mortar Data runs atop Amazon Web Services and currently supports Amazon S3 and MongoDB (hosted on Amazon EC2) as data sources.

%------------------------------%
Meter data management systems (MDMS):
Analytics-as-a-Service offerings
Advanced analytics as COTS software

