\subsection{Types of Factor Analysis}
There are two main types of extraction:
\begin{description}
\item[Principal components (PC):] Analyses all variance in the items, usually preferred when the goal is to reduce a set of variables down to a smaller number of factors and to create composite scores for these factors for use in subsequent analysis.
\item[Principal axis factoring (PAF):] Analyses shared variance amongst the items. Used more often for theoretical explorations of the underlying factor structure.
\end{description}
The most widely used method in factor analysis is the PAF method.

\subsection{Factor Loading}
The Factor loading is Correlation between a variable and a factor (or principal component), and the key to understanding the nature of a particular factor. Squared factor loadings indicate what percentage of the variance in an original variable is explained by a factor.

\subsection{Factor Rotation}
Factor rotation is the process of adjusting the factor axes to achieve a simpler and pragmatically more meaningful factor solution - the goal is a simple factor structure. The most commonly used type of rotation is \emph{\textbf{varimax}}. Others include equimax.

\textbf{Orthogonal factor rotation} is factor rotation such that their axes are maintained at 90 degrees (i.e. orthogonal). Each factor is independent of, or orthogonal to, all other factors. The correlation between the factors is determined to be zero. The opposite of this is \textbf{Oblique factor rotation}, rotation such that the extracted factors are allowed to be correlated
%-----------------------------------------------------------------------%
